---
title: "Heat Map in R"
author: "Nathan Byers"
date: "January 4, 2016"
output: html_document
---

This is a demonstration of how to make a heat map of Indiana. We'll use the 
`choroplethr` package, along with some other packages that will make life easier:

```{r, warning=FALSE, message=FALSE}
library(readxl)
library(dplyr)
library(choroplethr)
library(choroplethrMaps)
library(ggplot2)
```

In this demonstration we'll use data that is originally at the zip code level
and then scale it up to county and larger regions. Here we grab data about 
indiana zip codes from [www.stats.indiana.edu](http://www.stats.indiana.edu/):

```{r, eval=FALSE}
tmp = tempfile()
download.file(url = "http://www.stats.indiana.edu/maptools/county_zips.xls", 
              destfile = tmp)
countyZIP <- read_excel(tmp, sheet = 2, skip = 1)
names(countyZIP) <- c("FIPS", "county", "zip", "percent_zip_in_county")

# filter out a row of NAs at the end and drop the last column
countyZIP <- filter(countyZIP, !is.na(county))
countyZIP <- countyZIP[, -5]

# change class to data.frame
countyZIP <- as.data.frame(countyZIP)

# take a look see
head(countyZIP)
```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
load("county_zips.rda")
head(countyZIP)
```

So the `countyZIP` data frame has a record for every zip-county combination. If 
a zip code straddles two counties, then it will show up twice with the percentage
of area in each county recorded in the `percent_zip_in_county` column. For
simplicity sake we will assign one zip code to one county, based on the highest
percent of zip code area.

```{r}
countyZIP <- group_by(countyZIP, zip)
countyZIP <- filter(countyZIP, percent_zip_in_county == max(percent_zip_in_county))
```

```{r}
# 
# # categorize counties by region
# regions <- read.csv("counties.csv", stringsAsFactors = FALSE)
# countyZIPregions <- left_join(countyZIP, regions)
# 
# # get most recent numbers
# folders <- list.files("data")
# latest.folder <- max(folders)
# frequency <- read.csv(paste("data", latest.folder, "All Biobank Participant Zip Freq 6-22.csv", sep = "/"),
#                       stringsAsFactors = FALSE)
# countyZIPregions$ZIP.Code <- as.character(countyZIPregions$ZIP.Code)
# joined <- inner_join(frequency, countyZIPregions, by = c("Zip.Code" = "ZIP.Code"))
# 
# # summarize by region
# joined.region <- group_by(joined, Region)
# summed.region <- summarize(joined.region, total = sum(Participant.Frequency))
# 
# # attach region totals to county
# summed <- left_join(regions, summed.region)
# summed$County <- tolower(summed$County)
# 
# 
# data(county.regions)
# county.regions <- filter(county.regions, state.name == "indiana")
# 
# 
# summed <- merge(summed, county.regions, by.x = "County", by.y = "county.name")
# names(summed)[names(summed) == "total"] <- "value"
# # county_choropleth(summed,
# #                   #num_colors = 3,
# #                   legend     = "All IB",
# #                   state_zoom = c("indiana")) +  scale_fill_brewer(name = "All IB", palette="YlOrRd")
# county_choropleth(summed,
#                   legend     = "All IB",
#                   state_zoom = c("indiana")) +  
#   scale_fill_manual(
#     labels=c("0 to 299", "300 to 359", "360 to 399", 
#              "400 to 499", "500 to 699", "700 to 1,149",
#              "1,150 to 14,439"), 
#     values = c("#FFFFB2", "#FED976", "#FEB24C", "#FD8D3C",
#                "#FC4E2A", "#E31A1C", "#B10026"),
#     name = "All IB")
# ggsave(paste0("images/All_IB_regions", latest.folder, ".png"), width = 4, height = 4.5)
```
